---
title: "Introduction to densvis"
author:
  - name: Alan O'Callaghan
    email: alan.ocallaghan@outlook.com
package: densvis
output:
  BiocStyle::html_document:
    toc_float: yes
    fig_width: 10
    fig_height: 8
vignette: >
  %\VignetteIndexEntry{Introduction to densvis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    error = FALSE,
    warning=FALSE,
    message=FALSE,
    collapse = TRUE,
    comment = "#>"
)
library("BiocStyle")
```


# Introduction

# Setting up the data

We will illustrate the use of densvis
using simulated data.
We will first load the `densvis` and `Rtsne` libraries 
and set a random seed to ensure the t-SNE visualisation is reproducible
(note: it is good practice to ensure that a t-SNE embedding is robust
by running the algorithm multiple times).


```{r setup}
library("densvis")
library("Rtsne")
library("ggplot2")
theme_set(theme_bw())
set.seed(42)
```

```{r data}
data <- data.frame(
    x = c(rnorm(1000, 5), rnorm(1000, 0, 0.2)),
    y = c(rnorm(1000, 5), rnorm(1000, 0, 0.2)),
    class = c(rep("Class 1", 1000), rep("Class 2", 1000))
)
ggplot() +
    aes(data[, 1], data[, 2], colour = data$class) +
    geom_point(pch = 19) +
    scale_colour_discrete(name = "Cluster") +
    ggtitle("Original co-ordinates")
```

# Running t-SNE

The main functionality of the package lies in the `densne`
function. This function returns a matrix of t-SNE co-ordinates. In this case,
we pass in the 20 principal components computed based on the 
log-normalised counts. We colour points based on the discrete 
cell types identified by the authors.
We also set `dens_frac` (the fraction of optimisation steps that consider
the density preservation) and `dens_lambda` (the weight given to density
preservation relative to the standard t-SNE objective) each to 0.5.

```{r run}
fit1 <- densne(data[, 1:2], dens_frac = 0.5, dens_lambda = 0.5)
ggplot() +
    aes(fit1[, 1], fit1[, 2], colour = data$class) +
    geom_point(pch = 19) +
    scale_colour_discrete(name = "Class") +
    ggtitle("Density-preserving t-SNE") +
    labs(x = "t-SNE 1", y = "t-SNE 2")
```

If we run t-SNE on the same data, we can see that the density-preserving
objective better represents the density of the data, 

```{r run-tsne}
fit2 <- Rtsne(data[, 1:2])
ggplot() +
    aes(fit2$Y[, 1], fit2$Y[, 2], colour = data$class) +
    geom_point(pch = 19) +
    scale_colour_discrete(name = "Class") +
    ggtitle("Standard t-SNE") +
    labs(x = "t-SNE 1", y = "t-SNE 2")
```

cowplot::plot_grid(g1, g2)


# Session information {.unnumbered}

```{r}
sessionInfo()
```
